On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-04-12 manuscripts/PNAS/chazthm_pnas.tex
chaz

\significancestatement{Many naturally occurring signals (visual scenery, speech, EEG, etc.) lacking domain-specific formal models can nonetheless be usefully characterized as linear combinations of few elementary waveforms drawn from a large `dictionary'. We give general conditions guaranteeing when such dictionaries are uniquely and stably determined by data. %, justifying the application of this model as a constraint in blind source separation. The result may help explain the observed universality of emergent representations in sparse coding models of some natural phenomena.
%These results provide theoretical grounding for  the application of dictionary learning to recovering sparse signals from unknown noisy compressive measurements.
}

@@

\keywords{Sparse coding, dictionary learning, identifiability, compressive sensing}

\begin{abstract}
Dictionary learning for sparse linear coding has exposed underlying structure in many natural signals.
However, there are few universal theorems guaranteeing uniqueness of model estimation independent of implementation.
Here, we prove that for all diverse enough datasets generated from the sparse coding model, latent dictionaries and codes are uniquely determined up to measurement error.  Applications are given to data analysis, engineering, and neuroscience. 
\end{abstract}

>>>>

\significancestatement{Many naturally occurring signals (visual scenery, speech, EEG, etc.) lacking domain-specific formal models can nonetheless be usefully characterized as linear combinations of few elementary waveforms drawn from a large `dictionary'. We give general conditions guaranteeing when such dictionaries are uniquely and stably determined by data. The result justifies the use of this model as a constraint for blind source separation, and may help explain the observed universality of emergent representations in sparse coding models of some natural phenomena.
%These results provide theoretical grounding for  the application of dictionary learning to recovering sparse signals from unknown noisy compressive measurements.
}

@@

\keywords{Sparse coding, dictionary learning, matrix factorization, compressive sensing, inverse problems}

\begin{abstract}
Dictionary learning for sparse linear coding has exposed some characteristic properties of many natural signals.
However, universal theorems which guarantee the consistency of estimation in this model are lacking.
Here, we prove that for all diverse enough datasets generated from the sparse coding model, latent dictionaries and codes are uniquely and stably determined up to measurement error.  Applications are given to data analysis, engineering, and neuroscience. 
\end{abstract}
